---
title: "COD Reproducibility Report"
output:
  html_document:
    toc: true
    toc_float: true
---

#### Article ID: TvdWP
#### Article doi: http://doi.org/10.1016/j.cognition.2016.10.001
#### Pilot: Mika Asaba
#### Co-pilot: Tom Hardwicke  
#### Start date: 03/14/2017
#### End date: 

-------

#### Methods summary: 

Participants were given a hypothetical portfolio of 100 euros and were asked to distribute it between a risky option and a safe option in each monthly period (1-172). The safe option was a cash deposit account offering .25% rate of return each month. The risky option was the Spanish stock index fund which offered monthly rates of return (obtained from July 1999 to September 2013). Participants were in one of four conditions: "shock experience" (had to make investments across all 172 periods; experienced an initial decrease until period 40), "no-shock experience" (started at period 40), "shock description" (started at period 100 but viewed description of first 40 periods), and "no-shock description" (started at period 100 and did not view description of first 40 periods).

#### Target outcomes: 

The researchers were interested in how shock affected risk-taking behavior. The measure of risk taking **R** is defined as the **proportion of a person’s investment in the index fund (vs. the safe cash deposit)**.

1. Shock (Shock experience + Shock description) vs. No-Shock (No-shock experience + No-shock description): Shock vs. No-shock groups took similar financial risks (shock condition; Rs = 29.7%) as did participants who were unaware of the trend (no-shock condition; Rns = 32.5%), Rs–ns = −2.8%, 95% CI [−7.2%, 1.6%].

2. Shock description vs. No-shock description: Participants who learned from description took almost identical risk in the shock (Rsd = 37.1%) and no-shock conditions (Rnsd = 38.1%), Rsd–nsd = −1% [−7%, 5%]. 

3. Shock experience vs. No-shock experience: Participants in the shock experience condition took less risk (Rse = 22.4%) than did participants in the no-shock experience condition (Rnse = 26.9%). Although this difference in the average R may not indicate a true population difference, Rse–nse = −4.5% [−11%, 1.7%], 

4. Shock experience vs. Shock description: Participants who experienced the shock also took less risk than participants who learned about it from description, Rse–sd = −14.7% [−20%, −0.1%].

------

The chunk below sets up some formatting options for the R Markdown document:

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

## Step 1: Load packages

Some useful packages are being loaded below:

```{r}
library(tidyverse) # for data munging
library(knitr) # for kable table formating
library(haven) # import and export 'SPSS', 'Stata' and 'SAS' Files
library(readxl) # import excel files
library(CODreports) # custom report functions
library(forcats) #manipulating factors in data frames
```

## Step 2: Load data

```{r}
d <- read_tsv("data/data.txt")
```

## Step 3: Tidy data

```{r}
#factor subject and condition columns
str(d)
d$subject <- as_factor(d$subject)
d$condition <- as_factor(d$condition)
```

## Step 4: Run analysis

### Pre-processing

Convert to percentages:

```{r}
d$perc_stock <- d$perc_stock*100 # to match values that they report 
```

Look only at 'evaluation window':

```{r}
d <- subset(d,d$period >= 100) # only look at periods 100-172
```

```{r}
# participant level means for each of the conditions
d_avgs <- d %>%
  na.omit() %>%
  group_by(subject,condition) %>%
  summarise(mean_perc = mean(perc_stock))

#make separate data frames for each condition
d_ShockExp <- subset(d_avgs,d_avgs$condition == "Shock experience")
d_NoShockExp <- subset(d_avgs,d_avgs$condition == "No-shock experience")
d_ShockDesc <- subset(d_avgs,d_avgs$condition == "Shock description")
d_NoShockDesc <- subset(d_avgs,d_avgs$condition == "No-shock description")

#combine conditions by Shock (yes/no) and type of exposure (Experience/Description)
d_Shock <- full_join(d_ShockExp,d_ShockDesc)
d_NoShock <- full_join(d_NoShockExp,d_NoShockDesc)
d_Exp <- full_join(d_ShockExp,d_NoShockExp)
d_Desc <- full_join(d_ShockDesc,d_NoShockDesc)
```

### Descriptive statistics
```{r}
##### MEANS #####

mean_Shock <- mean(d_Shock$mean_perc) #matches value reported (29.7%)

mean_NoShock <- mean(d_NoShock$mean_perc) #matches value reported (32.5%)

mean_ShockDesc <- mean(d_ShockDesc$mean_perc)
#matches value reported (37.1%)

mean_NoShockDesc <- mean(d_NoShockDesc$mean_perc)
#matches value reported (38.1%)

mean_ShockExp <- mean(d_ShockExp$mean_perc) #matches value reported (22.4%)

mean_NoShockExp <- mean(d_NoShockExp$mean_perc) #matches value reported (26.9%)

mean_ShockDesc <- mean(d_ShockDesc$mean_perc)

##### Difference in Means #####

#1. Shock vs. No-Shock
diff_Shock_NoShock <- mean_Shock - mean_NoShock
#matches value reported (-2.8%)

#2. Shock Desc vs. No-Shock Desc
diff_Desc <- mean_ShockDesc - mean_NoShockDesc
#matches value reported (-1%)

#3. Shock Experience vs. No-Shock Experience
diff_Exp <- mean_ShockExp - mean_NoShockExp
#matches value reported (-4.5%)

#4. Shock Experience vs. Shock Description
diff_ShockExpDesc <- mean_ShockExp - mean_ShockDesc
#matches value reported (-14.7%)
```

### Inferential statistics

```{r}
##### CONFIDENCE INTERVALS #####

#1. Shock vs. No-Shock
df_1 <- 100 + 100 - 2
critical_t_1 <- qt(.975,df_1)

s_Shock <- sd(d_Shock$mean_perc)/sqrt(100)
s_Shock_sq <- s_Shock^2
s_NoShock <- sd(d_NoShock$mean_perc)/sqrt(100)
s_NoShock_sq <- s_NoShock^2
sp_1 <- sqrt(((99*s_Shock_sq) + (99*s_NoShock_sq))/df_1)

a_1 <- sqrt((1/100)+(1/100))

interval_upper_1 <- diff_Shock_NoShock + (critical_t_1*sp_1*a_1)
compareValues(reportedValue <- 1.6, obtainedValue = -2.186)

interval_lower_1 <- diff_Shock_NoShock - (critical_t_1*sp_1*a_1)
compareValues(reportedValue = -7.2, obtainedValue = -3.348)

#2. Shock Description vs. No-Shock Description
df_2 <- 50 + 50 - 2
critical_t_2 <- qt(.975,df_2)

sd_ShockDesc <- sd(d_ShockDesc$mean_perc)/sqrt(50)
sd_ShockDesc_sq <- sd_ShockDesc^2
sd_NoShockDesc <- sd(d_NoShockDesc$mean_perc)/sqrt(50)
sd_NoShockDesc_sq <- sd_NoShockDesc^2
sp_2 <- sqrt(((49*sd_ShockDesc_sq) + (49*sd_NoShockDesc_sq))/df_2)

a_2 <- sqrt((1/50)+(1/50))

interval_upper_2 <- diff_Desc + (critical_t_2*sp_2*a_2)
compareValues(reportedValue = 5, obtainedValue = .105)

interval_lower_2 <- diff_Desc - (critical_t_2*sp_2*a_2)
compareValues(reportedValue = -7, obtainedValue = -2.155)

#3. Shock Experience vs. No-Shock Experience
df_3 <- 50 + 50 - 2
critical_t_3 <- qt(.975,df_3)

sd_ShockExp <- sd(d_ShockExp$mean_perc)/sqrt(50)
sd_ShockExp_sq <- sd_ShockExp^2
sd_NoShockExp <- sd(d_NoShockExp$mean_perc)/sqrt(50)
sd_NoShockExp_sq <- sd_NoShockExp^2
sp_3 <- sqrt(((49*sd_ShockExp_sq) + (49*sd_NoShockExp_sq))/df_3)

a_3 <- sqrt((1/50)+(1/50))

interval_upper_3 <- diff_Exp + (critical_t_3*sp_3*a_3)
compareValues(reportedValue = 1.7, obtainedValue = -3.407)

interval_lower_3 <- diff_Exp - (critical_t_3*sp_3*a_3)
compareValues(reportedValue = -11, obtainedValue = -5.610)

#4. Shock Experience vs. Shock Description
df_4 <- 50 + 50 - 2
critical_t_4 <- qt(.975,df_4)

sd_ShockExp <- sd(d_ShockExp$mean_perc)/sqrt(50)
sd_ShockExp_sq <- sd_ShockExp^2
sd_ShockDesc <- sd(d_ShockDesc$mean_perc)/sqrt(50)
sd_ShockDesc_sq <- sd_ShockDesc^2

sp_4 <- sqrt(((49*sd_ShockExp_sq) + (49*sd_NoShockExp_sq))/df_4)

a_4 <- sqrt((1/50)+(1/50))

interval_upper_4 <- diff_ShockExpDesc + (critical_t_4*sp_4*a_4)
compareValues(reportedValue = -.1, obtainedValue = -13.585)

interval_lower_4 <- diff_ShockExpDesc - (critical_t_4*sp_4*a_4)
compareValues(reportedValue = -20, obtainedValue = -15.787)
```

## Step 5: Conclusion

```{r}
codReport(Report_Type = 'joint',
          Article_ID = 'TvdWP', 
          Insufficient_Information_Errors = 0,
          Decision_Errors = 0, 
          Major_Numerical_Errors = 8, 
          Minor_Numerical_Errors = 0)
```

Means and differences for each of the comparison conditions were consistent with the values that the authors reported. However, I had difficulty matching the 95% confidence intervals that were reported. I attempted to manually calculate the confidence intervals with the formula they printed in footnote 2 on p. 368, but the confidence intervals that I obtained were much smaller than theirs. I also tried running t-tests on their comparisons of interest, but the confidence intervals given in the t-test summary were generally slightly larger than those that they reported. 

When I ran compareValues on a few of the comparisons, "MATCH" was printed even though the two numbers seemed to be large enough to qualify as a Major Numerical Error. In the table above, I counted those that were reported as "MATCH" as Major Numerical Errors.

Information about the package versions used in this report:

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
